---
title: 'Title'
author : 'Team: Code Go Gold'
output: html_document
---
## Introduction

##### A research conducted by David Popenoe, a sociologist at Rutgers University, asserted that fifty percent of marriages will end in divorce. Actually, divorce is both very personal and all too common. But there are many myths about divorce. Therefore, in this markdown we mainly focus on the divorce rate in the United States, as well we would provide some simple tips or advice to the next generations. 


## Set the environment
```{r lib, message=F, include=T}
# add all the lib you use
library(rMaps)
library(data.table)
library("FactoMineR")
library("factoextra")
library("gplots")
library(sunburstR)
library(TraMineR)
library(pipeR)
library(RColorBrewer)
library(stringr)

```

## General Picture of the Divorce Rate
```{r code = ichoropleth, echo=F, message=F, include=F}
path = '/Users/pengfeiwang/Desktop/Fall2016-proj1-grp8/output/Divorce_Rate/Divorce_rate_total.csv'
rate = fread(path)
rate = subset(rate, Year != 2014)
r1 = ichoropleth(Divorce_Rate ~ State, ncuts =5, rate, animate = "Year", legend = FALSE)
r1$save(destfile = 'general.html')
```

```{r viz, echo=FALSE}
# htmltools::includeHTML("file:///Users/pengfeiwang/Desktop/Fall2016-proj1-grp8/figs/general.html")
```

###### This choropleth encodes divorce rates per 1000 population for successive 12-monthe period from 2008 with a quantize scale ranging from 0 to 11%. As we can see, the divorce rate is continuously decreasing for almost half decade in every states. However, the absolute value remains very high.



###### The following part is a general summary of of the relationsip between marrige times and education levels. Here, education level is accounted as one's highest degree. In the first ballon fig, we can see that in every education level, those who has never divorced accunt for the largest group. Also, the trend are the same: population decreases as the divorce time increases. But one interesting thing is that the "never divorce" and "divorce once" group sizes in Low education and Master level are almost the same. However, the "divorce twice" and ">3 time divorce" groups in Low education level are much larger than they are in Master level.

###### For the sunburst fig, the innermost circle showes proportion of different divorce times, the middel circle shows the education level, and the outermost circle shows the major, whether it's STEM or not. As we can see, the never divorced group accounts for 64% of the total married people. For most of us, who are master of STEM students, stays in blue-light orange part. So we probably end up with never divorce or divorce once(never divorce 0.758%, divorce once 0.254%) but will never divorce more than twice

## viz1
```{r code}
# Chencheng's Part
#setwd("~/Documents/Courses/ADS/Project 1")
# data cleansing
colstokeep_pus <- c("AGEP", "MSP", "MARHT", "SCHL", "FOD1P", "SCIENGRLP")
dataset_a <- fread('csv_pus/ss14pusa.csv', select= colstokeep_pus)
dataset_b <- fread('csv_pus/ss14pusb.csv', select= colstokeep_pus)
data.fm<- data.frame(rbind(dataset_a, dataset_b)[AGEP >18,])
data.fm <- data.fm[!is.na(data.fm[,"MARHT"]),]
data.fm$DIV=data.fm$MARHT-1
data.fm$DIV[data.fm$MSP==4]=data.fm$DIV[data.fm$MSP==4]+1
data.fm[is.na(data.fm)] <- 0

#####################################################
# education level calculation
lowedu= subset(data.fm, SCHL<=15) #lower than high school
highschool= subset(data.fm, SCHL>= 16 & SCHL<= 20) #including high school, GED, Associate's degree or enter the college without degree
bachelor= subset(data.fm, SCHL==21) 
master= subset(data.fm, SCHL==22| SCHL ==23) #including master and professional degree
phd= subset(data.fm, SCHL==24)
data_edu= list(lowedu,highschool,bachelor, master, phd)

#####################################################
#generate marry-education matrix
div_edu= matrix(nrow=5,ncol=4)
rownames(div_edu) <- c("Low education", "High school", "Bachelor", "Master", "PhD")
colnames(div_edu) <- c("Never", "Once", "Twice", ">3 times")
for(i in 1:5){
  for(j in 1:4){
    div_edu[i,j]= nrow(subset(data_edu[[i]],DIV == j-1))
  }
}

#####################################################
#draw the balloonplot
balloonplot(t(as.table(div_edu)), main ="Divorce and Education level", xlab ="Divorce", ylab="Education level",
             dotsize = 8, text.size= 0.7, label = F, show.margins = FALSE)
#####################################################
#draw the sunburst plot

edu= subset(data.fm, select= c("DIV","SCHL","SCIENGRLP"))
edu[,1]= sapply(edu[,1], paste, "time(s)")
edu_test= edu ## mark ##
for(i in 1:nrow(edu_test)){
  if(edu_test[i,2]<15){
    edu_test[i,2]="Lowedu"
  }else if(edu_test[i,2]>=16 & edu_test[i,2]<=20){
    edu_test[i,2]="Highschool"
  }else if(edu_test[i,2]==21){
    edu_test[i,2]="Bachelor"
  }else if(edu_test[i,2]==22|edu_test[i,2]==23){
    edu_test[i,2]="Master"
  }else{
    edu_test[i,2]="PhD"
  }
}

#prepare for the sunburst plot
edu.seq <- seqdef(edu_test)
sun_edu=seqtab(edu.seq, tlim = 0, format = "STS" )
name=names(attributes(as.list(sun_edu))$weights)
freq=as.numeric(attributes(as.list(sun_edu))$freq[,2])
sun_edu.fm=data.frame(name,freq)
sun_edu.fm[,1]= str_replace_all(sun_edu.fm[,1],"-0"," ")
sun_edu.fm[,1]= str_replace_all(sun_edu.fm[,1],"-1","-STEM")
sun_edu.fm[,1]= str_replace_all(sun_edu.fm[,1],"-2","-Non_STEM")

cols=c(brewer.pal(6,"Set2"),brewer.pal(6,"Set3"))
sunburst(sun_edu.fm,colors=cols)

```

### 
```{r viz}

```



## viz2
```{r}
# Minghao's Part

```



## viz3
```{r}
# Kaisheng's Part

```

